---
title: "3.6.3 - 3.6.6"
output: html_document
---

3.6.3 Multiple Linear Regression
In order to fit a multiple linear regression model using least squares, we
again use the lm() function. The syntax lm(y ∼ x1 + x2 + x3) is used to
fit a model with three predictors, x1, x2, and x3. The summary() function
now outputs the regression coefficients for all the predictors.

```{r}
library(ISLR2)
lm.fit <- lm(medv ~ lstat + age, data = Boston)
summary(lm.fit)
```

The Boston data set contains 12 variables, and so it would be cumbersome
to have to type all of these in order to perform a regression using all of the
predictors. Instead, we can use the following short-hand:

```{r}
lm.fit <-lm(medv ~ ., data = Boston)
summary(lm.fit)
```

We can access the individual components of a summary object by name
(type ?summary.lm to see what is available). Hence summary(lm.fit)$r.sq
gives us the R2, and summary(lm.fit)$sigma gives us the RSE. The vif()
function, part of the car package, can be used to compute variance inflation
factors. Most VIF’s are low to moderate for this data. The car package is
not part of the base R installation so it must be downloaded the first time
you use it via the install.packages() function in R.

```{r}
library(car)
vif(lm.fit)
```

What if we would like to perform a regression using all of the variables but
one? For example, in the above regression output, age has a high p-value.
So we may wish to run a regression excluding this predictor. The following
syntax results in a regression using all predictors except age.

```{r}
lm.fit1 <-lm(medv ~.- age, data = Boston)
summary(lm.fit1)
```

Alternatively, the update() function can be used.

```{r}
lm.fit1 <-update(lm.fit,~.- age)
```

reviewed 3.6.4-3.6.6