---
title: "citibike prediction model"
author: "mary mctarsney"
---

# Read in data
Use the trips_per_day.tsv file that has one row for each day, the number of trips taken on that day, and 
the minimum temperature on that day.

```{r}
trips_per_day <- read_tsv('trips_per_day.tsv')
```

# Splitting the data 
Split the data into randomly selected training, validation, and test sets, with 90% of the data for training and 
validating the model, and 10% for a final test set (to be used once and only once towards the end of this exercise). 
You can adapt the code from last week's complexity control notebook to do this. When comparing possible models, 
you can use a single validation fold or k-fold cross-validation if you'd like a more robust estimate.
```{r}
set.seed(42)

num_days <- nrow(trips_per_day)
frac_train_validate <- 0.9
num_train_validate <- floor(num_days * frac_train_validate)

# randomly sample rows for the training set 
ndx <- sample(1:num_days, num_train_validate, replace=F)

# used to fit the model
trips_per_day_TEMP <- trips_per_day[ndx, ]

# used to evaluate the fit
trips_per_day_test <- trips_per_day[-ndx, ]
     
```