---
title: "Predicting Daily Citibike Trips"
author: "mary mctarsney"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, message=FALSE}
library(tidyverse)
library(modelr)
```

## Load data

```{r}
trips_per_day <- read_tsv("trips_per_day.tsv")
```

## Train/test split

Hold out 10% of the data as a final test set. Use the remaining 90% for
training and validation.

```{r}
  set.seed(42)
  num_days <- nrow(trips_per_day)
  frac_tv <- 0.9
  num_tv <- floor(num_days * frac_tv)

  # randomly sample rows for the training/validation set 
  ndx <- sample(1:num_days, num_tv, replace=F)

  trips_per_day_tv <- trips_per_day[ndx, ]

  # used for testing
  trips_per_day_test <- trips_per_day[-ndx, ]

  # now test data is set aside ->
  # from training/validation data, split 90% for training, 10% from validation
  num_days <- nrow(trips_per_day_tv)
  frac_train <- 0.9
  num_train <- floor(num_days * frac_train)

  # randomly sample rows for the training set 
  ndx <- sample(1:num_days, num_train, replace=F)

  #used for training
  trips_per_day_train <- trips_per_day_tv[ndx, ]

  # used for validation
  trips_per_day_validate <- trips_per_day_tv[-ndx, ] 

  # have trips_per_day_train, trips_per_day_validate, and trips_per_day_test
```

## Baseline model: temperature only

Start with the polynomial model from the complexity control notebook.

```{r}
# TODO: fit a linear model using poly(tmin, degree) on training data
# TODO: try different polynomial degrees and evaluate RMSE on validation data
K <- 1:8
train_err <- c()
validate_err <- c()
for (k in K) {
  
    # fit on the training data
    model <- lm(num_trips ~ poly(tmin, k, raw = T), data=trips_per_day_train)
    
    # evaluate on the training data
    train_err[k] <- sqrt(mean((predict(model, trips_per_day_train) - trips_per_day_train$num_trips)^2))

    # evaluate on the validate data
    validate_err[k] <- sqrt(mean((predict(model, trips_per_day_validate) - trips_per_day_validate$num_trips)^2))
}

plot_data <- data.frame(K, train_err, validate_err) %>%
  gather("split", "error", -K)

ggplot(plot_data, aes(x=K, y=error, color=split)) +
  geom_line() +
  scale_x_continuous(breaks=K) +
  xlab('Polynomial Degree') +
  ylab('RMSE')
```

## Feature engineering

Add features that might help: day of week, month, precipitation, etc.
Remember: only use information available *before* the day in question.

```{r}
trips_per_day <- trips_per_day %>%
  mutate(is_weekend = wday(ymd) %in% c(1,7), temp_difference=tmax-tmin, month=month(ymd), day_of_week=day(ymd))

# splitting data again with new features
  set.seed(50)
  num_days <- nrow(trips_per_day)
  frac_tv <- 0.9
  num_tv <- floor(num_days * frac_tv)

  # randomly sample rows for the training/validation set 
  ndx <- sample(1:num_days, num_tv, replace=F)

  trips_per_day_tv <- trips_per_day[ndx, ]

  # used for testing
  trips_per_day_test <- trips_per_day[-ndx, ]

  # now test data is set aside ->
  # from training/validation data, split 90% for training, 10% from validation
  num_days <- nrow(trips_per_day_tv)
  frac_train <- 0.9
  num_train <- floor(num_days * frac_train)

  # randomly sample rows for the training set 
  ndx <- sample(1:num_days, num_train, replace=F)

  #used for training
  trips_per_day_train <- trips_per_day_tv[ndx, ]

  # used for validation
  trips_per_day_validate <- trips_per_day_tv[-ndx, ] 

  # have trips_per_day_train, trips_per_day_validate, and trips_per_day_test
```

## Model comparison

Try different models and compare their validation performance.

```{r}
model1 <- lm(num_trips ~ I(tmin^4) + I(tmax^4) + prcp + snwd*snow + is_weekend + month + day_of_week, data = trips_per_day_train)

model2 <- lm(num_trips ~ I(tmin^4)*I(tmax^4) + prcp*snow + month + day_of_week*is_weekend, data=trips_per_day_train)

model3 <- lm(num_trips ~ I(tmin^4)*I(tmax^3) + prcp*tmin + snow*tmin + month + day_of_week*is_weekend, data=trips_per_day_train)

model4 <- lm(num_trips ~ I(tmin^4)*I(tmax^4) + prcp*tmin + snow*tmin + month*day_of_week + day_of_week*is_weekend, data=trips_per_day_train)

model5 <- lm(num_trips ~ I(tmin^4)*I(tmax^4) + prcp*tmin + snow*tmin + month*is_weekend, data=trips_per_day_train)

models <- list(model1, model2, model3, model4, model5)
for (model in models){
  pred <- trips_per_day_validate %>%
    add_predictions(model)
  
  pred %>%
   summarize(rmse = sqrt(mean((pred - num_trips)^2)),
            cor = cor(pred, num_trips),
            cor_sq = cor^2) %>%
    print
}

# MODEL1
# rmse   cor cor_sq
# 5518. 0.848  0.718

# MODEL2
#   rmse   cor cor_sq
# 5645. 0.851  0.725

# MODEL3
#   rmse   cor cor_sq
# 4860. 0.881  0.776

# MODEL4
#   rmse   cor cor_sq
# 4800. 0.880  0.774

# MODEL5
#   rmse   cor cor_sq
# 4736. 0.886  0.785
```


## Final model

Pick your best model and plot the results.

```{r}
# model5 performed the best 
final_model <- model5

pred <- trips_per_day_validate %>%
    add_predictions(final_model)

pred %>%
  ggplot(aes(x=ymd)) +
    geom_line(aes(y=pred)) +
    geom_point(aes(y=num_trips))
```

```{r}
pred %>%
  ggplot(aes(x=pred,y=num_trips))+
    geom_point()
```

## Save model

```{r}
save(final_model, file="citibike_model.RData")
```

## Test set evaluation

**Do this only once, at the very end!**

```{r}
# TODO: load your saved model and predict on trips_test
# TODO: compute RMSE on test set
# TODO: how do you think this model will perform on future data?
load("citibike_model.RData")

rmse(final_model, trips_per_day_test)
# 3381.09
rsquare(final_model, trips_per_day_test)
# 0.844865
```
I think these scores indicate that this model could perform reasonably well on future data, however I do think
the performance on future data wouldn't be as good (especially based on the chart demonstrating the fit of the model).