---
title: "Predicting Daily Citibike Trips"
author: "Your Name"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, message=FALSE}
library(tidyverse)
library(modelr)
```

## Load data

```{r}
trips_per_day <- read_tsv("trips_per_day.tsv")
```

## Train/test split

Hold out 10% of the data as a final test set. Use the remaining 90% for
training and validation.

```{r}
set.seed(42)

num_days <- nrow(trips_per_day)
frac_tv <- 0.9
num_tv <- floor(num_days * frac_tv)

# randomly sample rows for the training/validation set 
ndx <- sample(1:num_days, num_tv, replace=F)

trips_per_day_tv <- trips_per_day[ndx, ]

# used for testing
trips_per_day_test <- trips_per_day[-ndx, ]

num_days <- nrow(trips_per_day_tv)
frac_tv <- 0.9
num_tv <- floor(num_days * frac_tv)

# randomly sample rows for the training/validation set 
ndx <- sample(1:num_days, num_tv, replace=F)

trips_per_day_train <- trips_per_day_tv[ndx, ]

# used for testing
trips_per_day_validate <- trips_per_day_tv[-ndx, ]  
```

## Baseline model: temperature only

Start with the polynomial model from the complexity control notebook.

```{r}
# TODO: fit a linear model using poly(tmin, degree) on training data
# TODO: try different polynomial degrees and evaluate RMSE on validation data
K <- 1:8
train_err <- c()
validate_err <- c()
for (k in K) {
  
    # fit on the training data
    model <- lm(num_trips ~ poly(tmin, k, raw = T), data=trips_per_day_train)
    
    # evaluate on the training data
    train_err[k] <- sqrt(mean((predict(model, trips_per_day_train) - trips_per_day_train$num_trips)^2))

    # evaluate on the validate data
    validate_err[k] <- sqrt(mean((predict(model, trips_per_day_validate) - trips_per_day_validate$num_trips)^2))
}


plot_data <- data.frame(K, train_err, validate_err) %>%
  gather("split", "error", -K)

ggplot(plot_data, aes(x=K, y=error, color=split)) +
  geom_line() +
  scale_x_continuous(breaks=K) +
  xlab('Polynomial Degree') +
  ylab('RMSE')
```

## Feature engineering

Add features that might help: day of week, month, precipitation, etc.
Remember: only use information available *before* the day in question.

```{r}
K <- 1:8
train_err <- c()
validate_err <- c()
for (k in K) {
  
    # fit on the training data
    model <- lm(num_trips ~ I(tmin*tmax^4) + prcp + snow + tmin*snow + prcp*snow, data=trips_per_day_train)
    
    # evaluate on the training data
    train_err[k] <- sqrt(mean((predict(model, trips_per_day_train) - trips_per_day_train$num_trips)^2))

    # evaluate on the validate data
    validate_err[k] <- sqrt(mean((predict(model, trips_per_day_validate) - trips_per_day_validate$num_trips)^2))
}


plot_data <- data.frame(K, train_err, validate_err) %>%
  gather("split", "error", -K)

ggplot(plot_data, aes(x=K, y=error, color=split)) +
  geom_line() +
  scale_x_continuous(breaks=K) +
  xlab('Polynomial Degree') +
  ylab('RMSE')
```

## Model comparison

Try different models and compare their validation performance.

```{r}

```

## Final model

Pick your best model and plot the results.

```{r}

```

## Save model

```{r}

```

## Test set evaluation

**Do this only once, at the very end!**

```{r}
# TODO: load your saved model and predict on trips_test
# TODO: compute RMSE on test set
# TODO: how do you think this model will perform on future data?
```
